Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-07-15 paper/paper.tex
reenable plots that I should not have disabled Tested everything on bennet.
added quick explanation of stochastic to paper Tested everything on crystal.

\begin{multline}
  \frac{\partial \rho_{D:ADP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ADP}-k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  +\delta(d_w)k_{de}\sigma_{DE},\hspace{3.4cm}
\end{multline}
\begin{multline}
  \frac{\partial \rho_{D:ATP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ATP}+k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  -\delta(d_w)[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
\end{multline}
\begin{multline}
  \frac{\partial \rho_E}{\partial t} = \mathcal{D}_E\nabla^2\rho_E+\delta(d_w)k_{de}\sigma_{DE}
  -\delta(d_w)k_E \sigma_D \rho_E
\end{multline}
\begin{multline}
  \frac{\partial \sigma_d}{\partial t} = -k_E\sigma_D\rho_E
  +[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
  \label{eq:d-on-wall}
\end{multline}
\begin{multline}
  \frac{\partial \sigma_{DE}}{\partial t} = -k_{de}\sigma_{DE}+k_E\sigma_D\rho_E\hspace{3cm}
  \label{eq:FifthPDE}
\end{multline}

@@

Our diffusion and reaction rates are shown below.  We are interested
primarily in the effect of cellular size and shape on the protein
oscillations, so we follow Huang\cite{huang2003dynamic} and do not
deviate from the wild type values used in the cited work (see values
below).  Huang's simulations use total MinD and MinE concentrations of
$1,000/\mu m$ and $350/\mu m$, respectively, in a cylindrical cell of
radius $0.5\mu m$, and in our (non-cylindrical) cells we use the same
number of proteins per unit volume.  These concentration values are
1273 $\mu m^{-3}$ and 446 $\mu m^{-3}$, respectively.
\begin{gather*}
  \mathcal{D}_D = \mathcal{D}_{E} = 2.5\micron^2/\text{sec}\\
  k_D^{\textrm{ADP $\rightarrow$ ATP}} = 1/\textrm{sec,  }
  k_D = 0.025 \micron /\textrm{sec}\\
  k_{dD} = 0.0015 \micron^3/ \textrm{sec,  }
  k_{de} = 0.7/\textrm{sec}\\
  k_E = 0.093 \micron^3 /\textrm{sec}.
\end{gather*}
A 3D grid is then constructed in cartesian coordinates, with a grid
spacing of .05\micron. We define our cell shapes and solve the
reaction-diffusion partial differential equations numerically to
observe the time evolution of the MinD and MinE concentrations inside
the cell.

@@

\section{Stability Analysis and Mannik's Experimental Shapes}
In general we have found that the protein reaction sequence discussed
above results in extremely robust oscillatory behavoir of polar
selection and oscillation in a wide variety of cell shapes.  We have
also found that while all sizes of cell will in time settle into
regular oscillations, there are size dependent distinctions in the
number and placement of the spatial positions of maximization.
Regular oscillatory behavoir between two poles at opposite ends of the
longer cellular direction is found in cells which are long enough in
one dimension for the reaction chain to be unstable but short enough
in the perpendicular dimension to be stable.  The limits are seen when
both dimensions are too short to be unstable, so that initial
assymetry in protein concentration relaxes into a motionless steady
state (see Figure~\ref{box-mannik}, or when both are too great, so
that the proteins exhibit more than two poles and those at extreme
postitions in both dimensions.

@@

\begin{figure}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  %\includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
  %\includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  \caption{Box plots of Mannik shapes for sizes at the lower stability
    threshhold.  The vertical axis shows stacked and changing over
    time the total number of proteins that are of four different
    compound stages and in the different sections of the cell.  Both
    shapes have longer axis of roughly 2.2$\mu m$.}
  \label{box-mannik}
\end{figure}

>>>>

\begin{multline}
  \frac{\partial \rho_{D:ADP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ADP}-k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  +\delta(d_w)k_{de}\sigma_{DE},\hspace{3.4cm}
\end{multline}
\begin{multline}
  \frac{\partial \rho_{D:ATP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ATP}+k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  -\delta(d_w)[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
\end{multline}
\begin{multline}
  \frac{\partial \rho_E}{\partial t} = \mathcal{D}_E\nabla^2\rho_E+\delta(d_w)k_{de}\sigma_{DE}
  -\delta(d_w)k_E \sigma_D \rho_E
\end{multline}
\begin{multline}
  \frac{\partial \sigma_D}{\partial t} = -k_E\sigma_D\rho_E
  +[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
  \label{eq:d-on-wall}
\end{multline}
\begin{multline}
  \frac{\partial \sigma_{DE}}{\partial t} = -k_{de}\sigma_{DE}+k_E\sigma_D\rho_E\hspace{3cm}
  \label{eq:FifthPDE}
\end{multline}

@@

Our diffusion and reaction rates are shown below.  We are interested
primarily in the effect of cellular size and shape on the protein
oscillations, so we follow Huang\cite{huang2003dynamic} and do not
deviate from the wild type values used in the cited work (see values
below).  Huang's simulations use total MinD and MinE concentrations of
$1,000/\mu m$ and $350/\mu m$, respectively, in a cylindrical cell of
radius $0.5\mu m$, and in our (non-cylindrical) cells we use the same
number of proteins per unit volume.  These concentration values are
1273 $\mu m^{-3}$ and 446 $\mu m^{-3}$, respectively.
\begin{gather*}
  \mathcal{D}_D = \mathcal{D}_{E} = 2.5\micron^2/\text{sec}\\
  k_D^{\textrm{ADP $\rightarrow$ ATP}} = 1/\textrm{sec,  }
  k_D = 0.025 \micron /\textrm{sec}\\
  k_{dD} = 0.0015 \micron^3/ \textrm{sec,  }
  k_{de} = 0.7/\textrm{sec}\\
  k_E = 0.093 \micron^3 /\textrm{sec}.
\end{gather*}
A 3D grid is then constructed in cartesian coordinates, with a grid
spacing of .05\micron.

In order to further our understanding of the results of various
simulation approaches to this system, we have performed both a basic
finite analysis, deterministic simulation that is spatially and
temporally discrete, and a stochastic analysis that is spatially
discrete but continuous in time.  Our stochastic simulation method
follows the work of Kraus~\cite{kraus1996crosstalk} which in turn
follows a method introduced by Gillespie~\cite{gillespie1977exact}.

@@

\section{Stability Analysis and Mannik's Experimental Shapes}
In general we have found that the protein reaction sequence discussed
above results in extremely robust oscillatory behavoir of polar
selection and oscillation in a wide variety of cell shapes.  We have
also found that while all sizes of cell will in time settle into
regular oscillations, there are size dependent distinctions in the
number and placement of the spatial positions of maximization.
Regular oscillatory behavoir between two poles at opposite ends of the
longer cellular direction is found in cells which are long enough in
one dimension for the reaction chain to be unstable but short enough
in the perpendicular dimension to be stable.  The limits are seen when
both dimensions are too short to be unstable, so that initial
assymetry in protein concentration relaxes into a motionless steady
state (see Figure~\ref{box-mannik}), or when both are too great, so
that the proteins exhibit more than two poles and those at extreme
postitions in both dimensions.

@@

\begin{figure}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  \caption{Box plots of Mannik shapes for sizes at the lower stability
    threshhold.  The vertical axis shows stacked and changing over
    time the total number of proteins that are of four different
    compound stages and in the different sections of the cell.  Both
    shapes have longer axis of roughly 2.2$\mu m$.}
  \label{box-mannik}
\end{figure}
